---
title: "Random Forest"
author: "Etibar Aliyev"
date: "4/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Packages**   
```{r cars}
library("dplyr")
library("naniar")
library(magrittr)
library("car")
library("caret")
library("tidyverse")
library(broom)
theme_set(theme_classic())
library(pROC)
library("ROCR")
library(mlbench)
library(caret)
library(randomForest)
require(caTools)
```

**Read data**    
```{r}
diabetic <- read.csv(file="diabetic.csv", header=T, na.strings = "?")
diabetic2 <- subset(diabetic, select=-c(weight,payer_code))
```


#Early readmission variable
```{r}
diabetic2 %<>%
mutate(early.readmit=case_when(
 readmitted  %in% "<30" ~ 1,
 readmitted %in% c(">30","NO") ~ 0 ))
```

#Create new HbA1c variable
```{r}

diabetic2$HbA1c <- ifelse(diabetic$A1Cresult == "None","None",
ifelse(diabetic$A1Cresult == '>8' & diabetic2$change == 'Ch','High-change',      ifelse(diabetic$A1Cresult == '>8' & diabetic2$change == 'No','High-no change',   ifelse(diabetic$A1Cresult == '>7' | diabetic2$A1Cresult == 'Norm','Normal',0))))
```


#Gender
```{r}
diabetic2 %<>% replace_with_na(replace = list(gender = "Unknown/Invalid"))

```



#Discharge disposition
```{r}

diabetic2 %<>%
mutate(discharge_disposition=case_when(
 discharge_disposition_id %in% 1 ~ "Discharged to home",
 discharge_disposition_id %in% 2:29 ~ "Otherwise" ))
```


#Admission source 
```{r}


diabetic2 %<>%
mutate(admission.source=case_when(
 admission_source_id %in% c(7) ~ "Emergency",
 admission_source_id %in% c(1,2) ~ "Physician/Clinic",
 admission_source_id %in% c(3:6,8:26) ~ "Otherwise"
))

```



#Admitting physician specialty
```{r}

diabetic2$admit.spec <- 
ifelse(is.na(diabetic2$medical_specialty), "Unknown",
ifelse(diabetic2$medical_specialty == "InternalMedicine","IntMed",
ifelse(diabetic$medical_specialty == 'Cardiology','Card',      ifelse(startsWith(as.character(diabetic2$medical_specialty), "Surg"),'Surg',   ifelse(diabetic2$medical_specialty == 'Family/GeneralPractice','F-GP', 'Other')))))

```


#Primary diagnosis
```{r}

diabetic2$prim.diag <- ifelse(diabetic2$diag_1>=390 &                                                             diabetic2$diag_1<=459 |      
                              diabetic2$diag_1==785,   
                              "Circulatory",
                       ifelse(diabetic2$diag_1>=460 &  
                              diabetic2$diag_1<=519 |  
                              diabetic2$diag_1==786, 
                              "Respiratory",        
                       ifelse(diabetic2$diag_1>=520 &  
                              diabetic2$diag_1<=579 | 
                              diabetic2$diag_1==787, 
                              "Digestive",   
                       ifelse(diabetic2$diag_1>=250.0 & 
                              diabetic2$diag_1<251.0, 
                              "Diabetes",
                       ifelse(diabetic2$diag_1>=800 & 
                              diabetic2$diag_1<=999, 
                              "Injury",
                       ifelse(diabetic2$diag_1>=710 & 
                              diabetic2$diag_1<=739, 
                              "Musculoskeletal",      
                       ifelse(diabetic2$diag_1>=580 & 
                              diabetic2$diag_1<=629 | 
                              diabetic2$diag_1==788, 
                              "Genitourinary",          
                       ifelse(diabetic2$diag_1>=140 & 
                              diabetic2$diag_1<=239, 
                              "Neoplasms", "Other"))))))))
View(diabetic2$prim.diag)
```

#Race
```{r}

diabetic2 %<>%
mutate(race2=case_when(
 race %in% c("AfricanAmerican")~"AfricanAmerican" ,
 race %in% c("Caucasian")~"Caucasian" ,
 race %in% c("Asian","Hispanic","Other")~"Other" ,
 is.na(race)~"Missing"
))

```

#Age
```{r}
diabetic2 %<>%
mutate(age3=case_when(
 age %in% c("[0-10)","[10-20)","[20-30)")~"30 years old or younger" ,
 age %in% c("[30-40)","[40-50)","[50-60)")~"30 years old to 60 years old" ,
 age %in% c("[60-70)","[70-80)","[80-90)","[90-100)")~"60 years old or older" 
))


```


#Create numeric Age variable
```{r}
diabetic2 %<>%
mutate(age.num=case_when(
 age %in% c("[0-10)")~5 ,
 age %in% c("[10-20)")~15 ,
 age %in% c("[20-30)")~25 ,
 age %in% c("[30-40)")~35 ,
 age %in% c("[40-50)")~45 ,
 age %in% c("[50-60)")~55 ,
 age %in% c("[60-70)")~65 , 
 age %in% c("[70-80)")~75 ,
 age %in% c("[80-90)")~85 ,
 age %in% c("[90-100)")~95
))       

```

#Drop observations
Select first patient visit if multiple visits: Results in 71518 observations.
```{r}
diabetic.unique <- distinct(diabetic2, patient_nbr, .keep_all= TRUE)

```


#Remove patients discharged to hospice: Results in 70198 observations
```{r}
diabetic.unique2 <- subset(diabetic.unique, discharge_disposition_id != 13)
diabetic.unique3 <- subset(diabetic.unique2, discharge_disposition_id!=14)
```


#Remove patients who died: Results in 69973 observations.
```{r}
diabetic.unique4 <- subset(diabetic.unique3, discharge_disposition_id!=11)

diabetic.unique5 <- subset(diabetic.unique4, discharge_disposition_id!=19)

diabetic.unique6 <- subset(diabetic.unique5, discharge_disposition_id!=20)

diabetic.unique7 <- subset(diabetic.unique6, discharge_disposition_id!=21)

```


#Before running random forest transforming variables into factor and numeric variable as per na.action = na.roughfix requirement

```{r}

diabetic.unique7 <- transform(
  diabetic.unique7,
  early.readmit = as.factor(early.readmit),
  discharge_disposition_id = as.factor(discharge_disposition_id),
  admission_source_id =as.factor(admission_source_id),
  medical_specialty = as.factor(medical_specialty),
  age = as.factor(age),
  HbA1c=as.factor(HbA1c),
  diag_1 = as.numeric(diag_1),
  race = as.factor(race),
  admit.spec = as.factor(admit.spec),
  prim.diag = as.factor(prim.diag)
  )



sapply(diabetic.unique7, class)

```


#Creating train, test subsets
```{r}

sample = sample.split(diabetic.unique7$early.readmit, SplitRatio = .75)
train = subset(diabetic.unique7, sample == TRUE)
test  = subset(diabetic.unique7, sample == FALSE)

```

```{r}
dim(train)
dim(test)

```


#Running random forest and prediction
```{r}
rf <- randomForest(
  early.readmit ~ discharge_disposition_id +  admit.spec+ race + prim.diag+
          admission_source_id + time_in_hospital + age.num + HbA1c,data=train,na.action= na.roughfix
)

pred = predict(rf, newdata=test[-49])
cm = table(test[,49], pred)
print(cm)
```

```{r}
confusionMatrix(pred, test$early.readmit)
```
#Our accuracy is aroung 90% which is quite good result.

#Checking class error
```{r}

rf$confusion[, 'class.error']

```
#99% early readmitted

#Variable importance
```{r}
# estimate variable importance
importance <- varImp(rf, scale=FALSE)
# summarize importance
print(importance)

```

#Plotting variable importance
```{r}

varImpPlot(rf,type=2)

```
#Above given results and plot show that the most 3 important variables are time_in_hospital, discharge_disposition_id and prim diag.

In conclusion, as a result of comparison between 2 models random forest model proved to be quite robust in terms of results as well as fault-tolerant functionality. Considering the fact that K-fold cv method creates randomness it can also be used to compute results.   









